---
title: "3 : Homework"
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'homework3.html'))})
output: 
  html_document: rmdformats::readthedown
---
```{r, echo=F, message=F}
library(dplyr)
library(kableExtra) # highlights table when hover over a line
```

# Homework 4
1. **3.12 : Airlines sometimes overbook flights. Suppose that for a plane with 50 seats, 55 passengers have tickers. Define the random variable Y as the number of ticketed passengers who actually show up for the flight. The probability mass function of Y appears in the accompanying table.** 
```{r, echo=F}
y <- c(45, 46, 47, 48, 49, 50, 51, 52, 53, 54, 55)
p.y <- c(0.05, 0.10, 0.12, 0.14, 0.25, 0.17, 0.06, 0.05, 0.03, 0.02, 0.01)
df <- data.frame("y" = y, "p.y" = p.y)
df_new <- as.data.frame(t(df))
options(scipen = 50) # get rid of e's
rownames(df_new) <- c("y", "p(y)")
kable(df_new, col.names = NULL) %>%
  kable_paper() %>%
  kable_material(c( "hover"))
```
(a.) **What is the probability that the flight will accommodate all ticketed passengers who show up?**

P($y\leq50$)=(0.05+0.10+0.12+0.14+0.25+0.17)=0.83

(b.) **What is the probability that not all ticketed passengers who show up can be accommodated?**

P(y>50)=1-P($y\leq50$)=1-0.83=0.17

(c.) **If you are the first person on the standby list (which means you be the first one to get on the plane if there are any seats available after all ticketed passengers have been accommodated), what is the probability that you will be able to take the flight? What is this probability if you are the third person on the standby list?**

P(y<50)=(0.05+0.10+0.12+0.14+0.25)=0.66
P(y<48)=(0.05+0.10+0.12)=0.27

2. **3.22 : Refer to Exercise 13, and calculate and graph the cdf F(x). Then use it to calculate the probabilities of the events given in part (a), part (b), part (c), part (d) of that problem. **

**13. A mail-order computer business has six telephone lines. Let X denote the number of lines in use at a specified time. Suppose the pmf of X is as given in the accompanying table. **
```{r, echo=F}
x <- c(0:6)
p.x <- c(0.10, 0.15, 0.20, 0.25, 0.20, 0.06, 0.04)
df <- data.frame("x" = x, "p.x" = p.x)
df_new <- as.data.frame(t(df))
options(scipen = 50) # get rid of e's
rownames(df_new) <- c("x", "p(x)")
kable(df_new, col.names = NULL) %>%
  kable_paper() %>%
  kable_material(c( "hover"))
```
\usepackage{amsmath}
\[ 
F_x(x) = \begin{cases}
      0 & x < 0 \\
      0.10 & 0 \leq x < 1 \\
      0.25 & 1 \leq x < 2 \\
      0.45 & 2 \leq x < 3 \\
      0.70 & 3 \leq x < 4 \\
      0.90 & 4 \leq x < 5 \\
      0.96 & 5 \leq x < 6 \\
      1 & 6 \leq x 
   \end{cases}
\]
```{r, echo=F}
x <- c(0,0,1,2,3,4, 5, 6)
y <- c(0,0.1,.25,.45,.7,.9,.96,1)
lines <- data.frame("x" = x, "y" = y)
plot(lines, main = "cdf of F(x)",
     ylab = "Cummulative Distribution",
     xlab = "Lines in Use (at one time)")
```

(a.) **{at most three lines are in use}**

P(X$\leq$ 3)=0.70

(b.) **{fewer than three lines are in use}**

P(X<3)=0.45

(c.) **{at least three lines are in use}**

P(X$\geq$ 3)=1-P(X<3)=0.55

(d.) **{between two and five lines, inclusive, are in use}**

P(2$\leq$ X$\leq$ 5)=0.96-0.25=0.71

2. **3.24 : An insurance company offers its policyholders a number of different premium payment options. For a randomly selected policyholder, let X = the number of months between successive payments. The cdf of X is as follows: **

\[ 
F(x) = \begin{cases}
      0 & x < 1 \\
      0.30 & 1 \leq x < 3 \\
      0.40 & 3 \leq x < 4 \\
      0.45 & 4 \leq x < 6 \\
      0.60 & 6 \leq x < 12 \\
      1 & 12 \leq x 
   \end{cases}
\]

(a.) **What is the pmf of X?**

\[ 
p(x) = \begin{cases}
      0 & X \ne {1, 3, 4, 6, 12 }\\
      0.3 & X =  1\\
      0.1 & X =  3\\
      0.05 & X =  4\\
      0.15 & X = 6 \\
      0.40 & X =  12\\
   \end{cases}
\]

(b.) **Using just the cdf, compute P($3\leq X\leq 6$) and P($4\leq X$).**

P($3\leq X\leq 6$)=0.60-0.30=0.30
P($4\leq X$)=0.45

4. **3.30 : An individual who has automobile insurance from a certain company is randomly selected. Let Y be the number of moving violations for which the individual was cited during the last 3 years. The pmf of Y is** 
```{r, echo=F}
y <- c(0, 1, 2, 3)
p.y <- c(.6, .25, .1, .05)
df <- data.frame("y" = y, "p.y" = p.y)
df_new <- as.data.frame(t(df))
options(scipen = 50) # get rid of e's
rownames(df_new) <- c("y", "p(y)")
kable(df_new, col.names = NULL) %>%
  kable_paper() %>%
  kable_material(c( "hover"))
```

(a.) **Compute E(Y).**

E(Y)=$\sum_{all y}^{3} yp(x)$=$0\cdot 0.6+1\cdot 0.25+2\cdot 0.2+3\cdot 0.05$=0.6

(b.) **Suppose an individual with Y violations incurs a surcharge of 100$Y^2$ dollars. Calculate the expected amount of the surcharge.** 

E($100Y^2$)=100E($Y^2$)=100$\sum y^2p(y)$=100($0^2\cdot 0.6+1^2\cdot 0.25+2^2\cdot 0.2+3^2\cdot 0.05$)=100(1.1)=110

5. **3.35 : A small market orders copies of a certain magazine for its magazine rack each week. Let X = demand for the magazine, with pmf.**
```{r, echo=F}
x <- c(1:6)
p.x <- c("1/15", "2/15", "3/15", "4/15", "3/15", "2/15")
df <- data.frame("x" = x, "p.x" = p.x)
df_new <- as.data.frame(t(df))
options(scipen = 50) # get rid of e's
rownames(df_new) <- c("x", "p(x)")
kable(df_new, col.names = NULL) %>%
  kable_paper()%>%
  kable_material(c( "hover"))
```

**Suppose the store owner actually pays $2.00 for each copy of the magazine and the price to customers is $4.00. If magazines left at the end of the week have no salvage value, is it better to order three or four copies of the magazine? [Hint: For both three and four copies order, express net revenue as a function of demand X, and then compute the expected revenue.]**

Given the demand for the magazines is X, 
Let the profit be y, and define the number of magazines ordered to be n. Therefore, 
\[ 
y = \begin{cases}
      4x-2n & n > x \\
      2n & n \leq x \\
   \end{cases}
\]
```{r, echo=F}
x <- c(1:6)
y.3 <- c(-2, 2, 6, 6, 6, 6)
y.4 <- c(-4, 0, 4, 8, 8, 8)
df <- data.frame("x" = x, "y.3" = y.3, "y.4" = y.4)
df_new <- as.data.frame(t(df))
options(scipen = 50) # get rid of e's
rownames(df_new) <- c("x", "y when n = 3", "y when n = 4")
kable(df_new, col.names = NULL, row.names = ) %>%
  kable_paper() %>%
  kable_material(c( "hover"))
```

E[$y_3 (X)$]=$\sum_{i = 1}^{6} yp(x)$=(-2)($\frac{1}{15}$)+(2)($\frac{2}{15}$)+(6)($\frac{3}{15}$)+(6)($\frac{4}{15}$)+(6)($\frac{3}{15}$)+(6)($\frac{2}{15}$)=4.93

E[$y_4 (X)$]=$\sum_{i = 1}^{6} yp(x)$=(-4)($\frac{1}{15}$)+(0)($\frac{2}{15}$)+(4)($\frac{3}{15}$)+(8)($\frac{4}{15}$)+(8)($\frac{3}{15}$)+(8)($\frac{2}{15}$)=5.33

It is better to order 4 copies of magazines because the expected profit is greater than ordering 3 copies. 

6. **3.39 : A chemical supply company currently has in stock 100 lb of a certain chemical, which it sells to customers in 5lb batches. Let X = the number of batches ordered by a random chosen customer, and suppose that X has pmf.**
```{r, echo=F}
z <- c(1:4)
p.z <- c(.2, .4, .3, .1)
df <- data.frame("z" = z, "p.z" = p.z)
df_new <- as.data.frame(t(df))
options(scipen = 50) # get rid of e's
rownames(df_new) <- c("z", "p(z)")
kable(df_new, col.names = NULL) %>%
  kable_paper()%>%
  kable_material(c( "hover"))
```

**Compute E(X) and V(X). Then compute the expected number of pounds left after the next customer’s order is shipped and the variance of the number of pounds left. [Hint: The number of pounds left is a linear function of X.]**

E(X)=$\sum zp(z)$=(1)(0.2)+(2)(0.4)+(3)(0.3)+(4)(0.1)=2.3

E($X^2$)=$\sum z^2 p(z)$=(1)(0.2)+(4)(0.4)+(9)(0.3)+(16)(0.1)=6.1

V(X)=E($X^2$)-$[E(x)]^2$=6.1-$2.3^2$=0.81

Given X is the number of batches order,
Let y be the lb’s of a certain chemical currently in stock. 
Then, 

y = 100-5
```{r, echo=F}
z <- c(1:4)
y <- c(95, 90, 85, 80)
df <- data.frame("z" = z, "y" = y)
df_new <- as.data.frame(t(df))
options(scipen = 50) # get rid of e's
kable(df_new, col.names = NULL) %>%
  kable_paper()%>%
  kable_material(c( "hover"))
```

E(Y)=$\sum yp(z)$=(95)(0.2)+(90)(0.4)+(85)(0.3)+(80)(0.1)=88.5

E($Y^2$)=$\sum y^2 p(z)$=(9025)(0.2)+(8100)(0.4)+(7225)(0.3)+(6400)(0.1)=7852.5

V(Y) = E($Y^2$) - $[E(Y)]^2$ = 7852.5 - $88.5^2$ = 20.25

# Homework 5

coming soon . . . 